---
title: "Taylors maturity sampling data visualisation"
output: word_document

 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


* Set up a git repository for your analysis. 

* Tracking your work with git. 

* Make your data tidy.

* Explore the data through visualisation (min 5 plots).

* Develop a final image that tells the data's story.

* Make a KnitR report with your code and plots.

* Present to your classmates.



# Setting up working environments.
####Working directory and folders
In my DATASCHOOL working directory, I created a new folders to organise my work.

I have a folder for my raw data, scripts and figures.

Using the drop down in R studio I have also created files, these files are:

* Project in R studio

* Script files in R (I have multiple files for different task for the project)

* Markdown file



The advantage of creating a new project in R studio is that I am prompted to create a git repository

####Creating a GIT repository.
When I select create a new project in R studio, I go onto create a new directory and new project, here is a small checkbox 'create git repository'.
Once this is selected I can go to R studio terminal to double check that I have repository.

In the terminal 
$git status

#### Next step is to go to GitHub
log in, and create a new repository.

I selected public, and simply follow the instructions: 

'....or push on existing repository from command line'

copy the code and paste in R studio terminal - done!

This part is setting up a repository on Github which allows my work to be pushed from my local computer to Github.

####Staging, committing and pushing in R studio
Again R studio makes things easy.

Git pane has a list of my files.

Checkbox allows me to stage the files and then commit.

Adding a comment on the commit window.

At the end of the day, I can push my files to Github using the push icon in the git window.

#Importing data and make Tidy
####Import data

* Import tidyverse and ggplot2
(note tidyverse already had ggplot2)

* Import dplyer

* Import my data using read_csv

```{r import tidyverse}
library(tidyverse)
library(dplyr)

maturity_data <- read_csv("Raw_data/Taylors_maturity_2015_2016.csv")
```
####Check out what my data looks like
I have a few options here;

* head()

* str()

* View(maturity_data) 

* glimps
```{r look at data}
#str(maturity_data)
#head(maturity_data)
#View(maturity_data)
glimpse(maturity_data)
```
####Make my data tidy.
I have a number of problem with my data that needs to be fixed:

1. I have columns that need to be removed, ie I have number of columns that contain NA

2. I have column headings with white spaces

3. I don't have a unique ID column

4. My date column is not represented at a date rather a string.
  I should consider splitting this column out to day, month, year

Consider creating a new data column that has year as factor.

####1. Select data I want to use / remove the data I don't want.
Reduce my data set from 31 variable to 18.

```{r delete the columns I dont want to use}

mat_tidy1 <- select(maturity_data,c(1:18))
#glimpse(mat_tidy1)
```
I played around with a number of other option.
(I was not convinced it worked 
It looked like this worked when you run head but when you run str() you get the original list).

```{r select the data I want}
#test1_1 <- select(maturity, -starts_with("X"))
#head(test1_1)

```

####2. Remove the white spaces in heading names.
This can be done one heading at a time using this function.

names(maturity_data_clean)[names(maturity_data_clean) == 'sampling event'] <- 'sampling_event'

But much better way is to use this function.
Which replaces both the white spaces and / with _.
```{r removing the white spaces from the heading}
names(mat_tidy1) <-str_replace_all(names(mat_tidy1), c(" " = "_", "/" = "_" ))
#glimpse(mat_tidy1)
```

*Help I would like to make a new df with the new headings but I can't get this to work for me*



####3. Create a unique ID.

**3a Create a new data column for year as float**

```{r create a new data clm for year as float}
mat_tidy1 <- mat_tidy1 %>%
  mutate(Year_a = Year)

mat_tidy1$Year_a <- as.factor(mat_tidy1$Year_a)
#glimpse(mat_tidy1)
```

**3b create a new data coloum that combines Year and GPS_sampling__pt_ID**

```{r new data coloum with unquie ID}
mat_tidy3 <- mat_tidy1 %>%
  unite(ID, Year, GPS_sampling__pt_ID)

#glimpse(mat_tidy3)
```
Another option would be: This would keep input data columns.

unite(ID, Year, GPS_sampling__pt_ID, sep = "_", remove = FALSE)

  
####4. My date column is not represented at a date rather a string.
I should consider splitting this coloumn out to day, month, year.

```{r split out date clm}
mat_tidy4 <- mat_tidy3 %>%
  separate(Sampling_date, c("Day", "Month", "Year"), sep = '/')
mat_tidy4

#head(mat_tidy4)
```
These columns are character I want them to be integers. 
```{r make day month year as interger}
mat_tidy4$Day <- as.integer(mat_tidy4$Day)
mat_tidy4$Month <- as.integer(mat_tidy4$Month)
mat_tidy4$Year <- as.integer(mat_tidy4$Year)
#glimpse(mat_tidy4)
```

#Plotting data
###Plot grape variable vs time or degree days.
Make an empty plot

```{r make an empty plot before adding data}
#ggplot(data=mat_tidy4, aes(x = sampling_event, y = brix))
```

Now fill my plot with data.

```{r fill plot with sampling event vs brix}
ggplot(data=mat_tidy4, aes(x = sampling_event, y = brix, colour = Year_a))+
  geom_point(alpha = 0.3)
```

Each year had a number of sampling events which occurred through the season.

In 2015 we sampled 30 vines on 7 separate  events.

In 2016 we sampled 30 vines on 8 separate events.

Plotting by event makes the data points sit ontop of each other.
Plotting by date (on the x axis) will mean that there is a big gap between years.
I have calculated the growing degree days for each year, based on local weather data, this could be a way of displaying the x axis.

```{r fill plot with GFV vs brix}
ggplot(data=mat_tidy4, aes(x = GFV, y = brix, colour = Year_a))+
  geom_point(alpha = 0.3) +
  labs(x = "Degree days",y = "Brix",title = ("maturity of grapes"),colour = "Year") +
    theme(axis.text.x=element_text(), axis.ticks.x = element_blank())
```

Formatting...

```{r same as above with formatting}
ggplot(data=mat_tidy4, aes(x = GFV, y = brix, colour = Year_a))+
  geom_point(alpha = 0.3) +
  labs(x = "Degree days",y = "Brix",title = ("maturity of grapes"),colour = "Year") +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(), 
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
```

##Try all the grape properties...

```{r plot all the grape properties}

#X = GFV and Y = pH
ggplot(data=mat_tidy4, aes(x = GFV, y = pH, colour = Year_a))+
  geom_point(alpha = 0.3) +
  labs(x = "Degree days",y = "pH",title = ("maturity of grapes"),colour = "Year") +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(), 
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
        
#X = GFV and Y = TA
ggplot(data=mat_tidy4, aes(x = GFV, y = TA, colour = Year_a))+
  geom_point(alpha = 0.3) +
  labs(x = "Degree days",y = "TA",title = ("maturity of grapes"),colour = "Year") +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(), 
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))

#X = GFV and Y = bunch_wt
ggplot(data=mat_tidy4, aes(x = GFV, y = bunch_wt, colour = Year_a))+
  geom_point(alpha = 0.3) +
  labs(x = "Degree days",y = "Bunch weight",title = ("maturity of grapes"),colour = "Year") +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(), 
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
#X = GFV and Y = Mean_berry_wt
ggplot(data=mat_tidy4, aes(x = GFV, y = Mean_berry_wt, colour = Year_a))+
  geom_point(alpha = 0.3) +
  labs(x = "Degree days",y = "Mean berry weight",title = ("maturity of grapes"),colour = "Year") +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(), 
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
#X = GFV and Y = Colour_per_berry_Wt
ggplot(data=mat_tidy4, aes(x = GFV, y = Colour_per_berry_Wt, colour = Year_a))+
  geom_point(alpha = 0.3) +
  labs(x = "Degree days",y = "Colour per berry weight",title = ("maturity of grapes"),colour = "Year") +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(), 
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))

#X = GFV and Y = Phenolic_per_berry_Wt
ggplot(data=mat_tidy4, aes(x = GFV, y = Phenolic_per_berry_Wt, colour = Year_a))+
  geom_point(alpha = 0.3) +
  labs(x = "Degree days",y = "Phenolic per berry weight",title = ("maturity of grapes"),colour = "Year") +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(), 
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))

```

##Summary of grape variable vs degree days.
Most grape properties change as the degree days increase (brix, pH, TA and colour).
Handful of grape properties have a high amount of variation and you can't see change over season (bunch weight, berry weight and phenolics to a degree).

Does the location of the sampling in the vineyard matter?

#Do grape properties change depending on location.

Plotting variation through spaces vs variation through time.
Using brix and bunch weight as example.

##Plot brix for each sample point.
```{r plot by sample location}
ggplot(data=mat_tidy4, aes(x = Sample, y = brix, colour = Year_a))+
  geom_point(alpha = 0.5) + facet_wrap(~ sampling_event)+
labs(x = "sample location",y = "Brix",title = ("maturity of grapes"),colour = "Year") +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(), 
        axis.ticks.x = element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))

```

This plot is saying the early in the season differences between vineyard location and year is not as obvious as late in the year.
This is quite a broad generalisation.


## Is sample location associated with high brix values?


```{r}
mat_tidy4 %>%
    filter(Year_a == "2015") %>%
  ggplot(aes(x = POINT_X, y = POINT_Y, colour = brix)) +
  geom_point(size = 2.5) +
  facet_wrap(~ GFV) + 
  labs(                                      
    x = "Easting",
    y = "Northing",
    title = "2015",
    colour = "Brix") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))

mat_tidy4 %>%
  filter(Year_a == "2016") %>%
  ggplot(aes(x = POINT_X, y = POINT_Y, colour = brix)) +
  geom_point(size = 2.5) +
  facet_wrap(~ GFV) + 
  labs(                                      
    x = "Easting",
    y = "Northing",
    title = "2016",
    colour = "Brix") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
```

What I really want to see is each pane with its own scale, it is hard to remove the effect of time on these graphs.

Standardize brix per sampling event.


### Harvest sampling is pretty important.

```{r brix at harvest by location 2015 and 2016}
mat_tidy4 %>%
  filter(GFV == "3277") %>%
  ggplot(aes(x = POINT_X, y = POINT_Y, colour = brix)) +
  geom_point(size = 2.5) +
  labs(                                      
    x = "Easting",
    y = "Northing",
    title = "Harvest 2015",
    colour = "brix") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))

mat_tidy4 %>%
  filter(GFV == "3499") %>%
  ggplot(aes(x = POINT_X, y = POINT_Y, colour = brix)) +
  geom_point(size = 2.5) +
  labs(                                      
    x = "Easting",
    y = "Northing",
    title = "Harvest 2016",
    colour = "brix") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))

```

```{r late saeson sampling }
mat_tidy4 %>%
  filter(GFV > "3000") %>%
  ggplot(aes(x = POINT_X, y = POINT_Y, colour = brix, shape = Year_a)) +
  geom_point(size = 1.5, alpha = 0.5) +
  labs(                                      
    x = "Easting",
    y = "Northing",
    title = "Late season sampling",
    colour = "brix") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
```


#### Bunch weight


```{r Bunch weight for 2015 and 2016}
mat_tidy4 %>%
  filter(Year_a == "2015") %>%
  ggplot(aes(x = POINT_X, y = POINT_Y, colour = bunch_wt)) +
  geom_point(size = 2.5) +
  facet_wrap(~ GFV) + 
  labs(                                      
    x = "Easting",
    y = "Northing",
    title = "2015",
    colour = "Bunch weight") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))

mat_tidy4 %>%
  filter(Year_a == "2016") %>%
  ggplot(aes(x = POINT_X, y = POINT_Y, colour = bunch_wt)) +
  geom_point(size = 2.5) +
  facet_wrap(~ GFV) + 
  labs(                                      
    x = "Easting",
    y = "Northing",
    title = "2016",
    colour = "Bunch weight") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
```

##Summary of sample location vs grape properties.
Plotting just coordinated vs grape properties is not telling me much.

Need to overlay with something else, location in the vineyard is a surrogate for soil condition elevation etc…

I need to add another variable to my data set!
Or / and standardise my data per sampling event.


#Plotting grape variable against each other.


```{r Brix vs all most}

#Brix vs pH
mat_tidy4 %>%
  ggplot(aes(x = brix, y = pH, colour = GFV, shape = Year_a)) +
  geom_point(size = 2.5, alpha =0.8) +
  labs(                                      
    x = "Brix",
    y = "pH",
    title = "",
    colour = "Degree days", 
    shape = "Year") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_text(),
        axis.text.y=element_text(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
#Brix vs TA
mat_tidy4 %>%
  ggplot(aes(x = brix, y = TA, colour = GFV, shape = Year_a)) +
  geom_point(size = 2.0, alpha =0.8) +
  labs(                                      
    x = "Brix",
    y = "TA",
    title = "",
    colour = "Degree days", 
    shape = "Year") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_text(),
        axis.text.y=element_text(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
glimpse(mat_tidy4)

#Brix vs bunch_wt
mat_tidy4 %>%
  ggplot(aes(x = brix, y = bunch_wt, colour = GFV, shape = Year_a)) +
  geom_point(size = 2.0, alpha =0.8) +
  labs(                                      
    x = "Brix",
    y = "Bunch weight",
    title = "",
    colour = "Degree days", 
    shape = "Year") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_text(),
        axis.text.y=element_text(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
#Brix vs Mean_berry_wt
mat_tidy4 %>%
  ggplot(aes(x = brix, y = Mean_berry_wt, colour = GFV, shape = Year_a)) +
  geom_point(size = 2.0, alpha =0.8) +
  labs(                                      
    x = "Brix",
    y = "Mean berry weight",
    title = "",
    colour = "Degree days", 
    shape = "Year") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_text(),
        axis.text.y=element_text(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
#Brix vs Colour_per_berry_Wt
mat_tidy4 %>%
  ggplot(aes(x = brix, y = Colour_per_berry_Wt, colour = GFV, shape = Year_a)) +
  geom_point(size = 2.0, alpha =0.8) +
  labs(                                      
    x = "Brix",
    y = "Colour per berry weight",
    title = "",
    colour = "Degree days", 
    shape = "Year") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_text(),
        axis.text.y=element_text(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
#Phenolic_per_berry_Wt
mat_tidy4 %>%
  ggplot(aes(x = brix, y = Phenolic_per_berry_Wt, colour = GFV, shape = Year_a)) +
  geom_point(size = 2.0, alpha =0.8) +
  labs(                                      
    x = "Brix",
    y = "Phenolic per berry weight",
    title = "",
    colour = "Degree days", 
    shape = "Year") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_text(),
        axis.text.y=element_text(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
```

Do pH and TA have relationship?

```{r}
#pH vs TA
mat_tidy4 %>%
  ggplot(aes(x = pH, y = TA, colour = GFV, shape = Year_a)) +
  geom_point(size = 2.0, alpha =0.8) +
  labs(                                      
    x = "pH",
    y = "TA",
    title = "",
    colour = "Degree days", 
    shape = "Year") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_text(),
        axis.text.y=element_text(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))


#pH vs colour
mat_tidy4 %>%
  ggplot(aes(x = pH, y = Colour_per_berry_Wt, colour = GFV, shape = Year_a)) +
  geom_point(size = 2.0, alpha =0.8) +
  labs(                                      
    x = "pH",
    y = "Colour per berry Wt",
    title = "",
    colour = "Degree days", 
    shape = "Year") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_text(),
        axis.text.y=element_text(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
#phenolics vs colour
mat_tidy4 %>%
  ggplot(aes(x = Phenolic_per_berry_Wt, y = Colour_per_berry_Wt, colour = GFV, shape = Year_a)) +
  geom_point(size = 2.0, alpha =0.8) +
  labs(                                      
    x = "Phenolic per berry Wt",
    y = "Colour per berry Wt",
    title = "",
    colour = "Degree days", 
    shape = "Year") +
  theme(panel.grid = element_blank(), #remove panel lines
        axis.text.x=element_text(),
        axis.text.y=element_text(),
        panel.background = element_rect(fill = "white",
                                        colour = "black"))
```

##Summary of grape variable against each other.
Many grape properties are correlated.
Brix with pH, TA and colour
Can see a seasonal effects with phenolics especially early in the season.
