---
title: "Grape maturity data exploration"
author: "Jackie"
date: "7 August 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Add in table of contents when I work out how


# Setting up my working environmnets.
##Working directory and folders
In my DATASCHOOL working directory, I created a new folders to organise my work.
I have a folder for my raw data, scripts and figures.
Using the drop down in R studio I have also created files, which I will do my work:
These files are:
1.Create a new project in R studio
2.Create a new script file in R
3.Create a new markdown file

The advantage of creating a new project in R studio is that I am prompted to create a git repository

##Creating a GIT repository.
When I select create a new project in R studio, I go onto create a new directory and new project, here is a small checkbox 'create git repository'.
Once this is selcted I can go to R studio Termial to double check that I have respository.

```{r}
## Question can I add code in markdown that relates to terminal?
#> git status (this gives an error)
#$ git status (this gives an error)
```
## Next step is to go to GitHub
log in, and create a new repostitory.
I selecetd public 
Simply follow the instructions 
'....or push on existing repository from command line'
copy the code and paste in R studio terminal - done!

This part is setting up a repsoitory on Github which allows my work to be pushed from my local computer to Github - Safe and sound!

##Staging, commiting and pushing in R studio
Again R studio makes things easy for me!
Git pane has a list of my files
Checkbox allows me to stage the files and then commit.
Adding a comment on the commit window.

At the end of the day I can push my files to Github using the push icon in the git window.

##Getting ready for my data.
Import tidyverse and ggplot2
(note tidyverse already had ggplot2)

import my data using read_csv

```{r}
library(tidyverse)

maturity_data <- read_csv("Raw_data/Taylors_maturity_2015_2016.csv")
```

##Display my data as a summary in Rstudio
```{r}
head(maturity_data)
```
Or I can use

```{r}
str(maturity_data)

nrow(maturity_data)
ncol(maturity_data)
```


#I have empty columns of data that should be removed... how to do this is R?
I can do it one at a time using 
maturity_data$X30<- NULL

Or I can remove all the empty coloumns using this function, creating a new data frame called maturity_data_clean.

Check out if I have done it!
```{r}

maturity_data_clean <- maturity_data[,colSums(is.na(maturity_data))<nrow(maturity_data)]
head(maturity_data_clean)
```

Can be done a better way use tidy drop  

 

drop_na (maturity_data, X27)
I can do one clm at a time - but what about multiple?
drop_na () this won't work for me...I think this is just removing the NA from the rows


maturity1 <- select(maturity, -starts_with("X"))
```{r}
#maturity1 <- select(maturity, -starts_with("X"))
#head(maturity1)
```



 My heading names have gaps and R doesnt like this!
 I need to change heading names
 Again there is a better way to do this is tidyvesre... come back to this
```{r}
names(maturity_data_clean)[names(maturity_data_clean) == 'sampling event'] <- 'sampling_event'
```
I can go onto change all the others with spaces... better to set up file without space to start with.
This works... to relpace white spaces with names..
names(maturity2_a_rename)<-str_replace_all(names(maturity2_a_rename), c(" " = "." , "," = "" ))
head(maturity2_a_rename)
same as above but with _ not .
names(maturity1)<-str_replace_all(names(maturity1), c(" " = "_" ))
head(maturity1)

##Split my date cloumn into day month year, but this will overwrite my Sampling_date clm
```{r}
#data1 <- maturity1 %>%
  #separate(Sampling_date, c("Day", "Month", "Year"), sep = '/')
#data1
```

##Change Year to factors
```{r}
#play_year$Year <- as.factor(play_year$Year)
#head(play_year)
```


##Make may data tidy
I need a unquie identify that is not repeated in the data clm
```{r}

```

All is not lost beacuse I only really need year and I need to split out date and reformat it.

This works in my R studio...Yes but it removes the data clms I want
Need to make a copy of them first?
maturity_data_clean <- maturity_data_clean %>%
  unite(ID, Year, GPS_sampling_pt_ID)
str(maturity_data_clean)

```{r}

```

Ian method would be better
maturity_data_clean %>%
#  mutate(ID_1 = )

ggplot(data = maturity_data_clean, aes(x = Sampling_date, y = pH))

```{r}

```

