---
title: "Taylors maturity sampling data visulaiation"
output: github_document

 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#*Add in table of contents when I work out how*

* Set up Git

* import data and make tidy

* display data 


# Setting up working environmnets.
##Working directory and folders
In my DATASCHOOL working directory, I created a new folders to organise my work.

I have a folder for my raw data, scripts and figures.

Using the drop down in R studio I have also created files, in which I will do my work.

These files are:

* Create a new project in R studio

* Create a new script file in R

* Create a new markdown file



The advantage of creating a new project in R studio is that I am prompted to create a git repository

##Creating a GIT repository.
When I select create a new project in R studio, I go onto create a new directory and new project, here is a small checkbox 'create git repository'.
Once this is selcted I can go to R studio Termial to double check that I have respository.


In the terminal 
$git status
## Next step is to go to GitHub
log in, and create a new repostitory.

I selecetd public 

Simply follow the instructions 

'....or push on existing repository from command line'

copy the code and paste in R studio terminal - done!

This part is setting up a repsoitory on Github which allows my work to be pushed from my local computer to Github - Safe and sound!

##Staging, commiting and pushing in R studio
Again R studio makes things easy for me!

Git pane has a list of my files.

Checkbox allows me to stage the files and then commit.

Adding a comment on the commit window.

At the end of the day I can push my files to Github using the push icon in the git window.

#Importing data.

*Import tidyverse and ggplot2
(note tidyverse already had ggplot2)

*Import dplyer

*Import my data using read_csv

```{r import tidyverse}
library(tidyverse)
library(dplyr)

maturity_data <- read_csv("Raw_data/Taylors_maturity_2015_2016.csv")
```
#Check out what my data looks like
I have a few option here;

*head()

*str()

*use the table from the environment panel 

*glimps
```{r look at data}
str(maturity_data)
head(maturity_data)
```
#Make my data tidy.
I have a number of problem with my data that needs to be fixed:

1.I have coloumn that need to be removed that have NA

2.I have coloumn headings with white spaces

3.I don't have a unquie ID coloumn

4.My date coloumn is not represented at a date rather a string.
  I should consider splitting this coloumn out to day, month, year

5.Consider creating a new data coloumn that has year as factor.

##Select data I want to use / remove the data I don't want.

```{r delete the columns I dont want to use}

```

